---
title: "visulisation hate crime"
author: "Chaitanya Shekar"
output:
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(lessR)
library(tidyverse)
```
READ THE DATA
```{r}
hate_crime = read.csv("../../data/modified-data/cleaned_hate_crime_record_data.csv")
head(hate_crime)
```

DATA VISUALISATION
- pie chart for offender race
```{r}
offense_race_count = hate_crime %>%
  count(OFFENDER_RACE)
offense_race_count=offense_race_count[-1,]
hsize = 3
offense_race_count <- offense_race_count %>% 
  mutate(x = hsize)

ggplot(offense_race_count, aes(x = hsize, y = n, fill = OFFENDER_RACE)) +
  geom_col(color = "black") +
  coord_polar(theta = "y") +
  ggtitle("Distribution of Offender Race 1991-2020")+
  scale_fill_brewer(palette = "GnBu") +
  xlim(c(0.2, hsize + 0.5)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
```

-BAR GRAPH FOR MONTHWISE INCIDENTS

CODE:

hate_crime(INCIDENT_DATE) = as.Date(hate_crime(INCIDENT_DATE), "%m/%d/%Y")

month_incidents = hate_crime %>%

  group_by(month(INCIDENT_DATE)) %>%
  
  summarize(n=n())
  
month = c("Jan", "Feb","March","April","May","June","July","Aug","Sep","Oct","Nov","Dec")

month_incidents = cbind(month_incidents,month);month_incidents

barplot(height=month_incidents$n, names=month_incidents$month, col=rgb(0.8,0.1,0.1,0.6),xlab="Month", ylab="Number of Incidents", main="Crime incidents monthwise from 1991-2020", cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)

